---
title: "CS889 - To Clack or Not to Clack: An Investigation of Audio-Tactile cues for Typing Efficiency"
author: "Jia Wu (jr2wu)"
output: html_document
---

```
# Install packages.
install.packages("ez")
install.packages("dplyr")
install.packages("reshape")
install.packages("knitr")
install.packages("rcompanion")
```

```{r warning=FALSE, message=FALSE}
# Load the libraries
library(ez)
library(dplyr)
library(reshape)
library(knitr)
library(rcompanion)

# Load the data
data.dir <- "../data/unzipped"
file.name <- "summarized_data.csv"

file <- paste(data.dir,file.name,sep="/")
data <- read.csv(file)

# Clean the Data
data$X <- NULL
data$ID <- as.integer(data$ID) #TODO" Convert the participants to a numeric value
```

First we must examine the data. We examine the normality of each data distribution for **WPM** first.

```{r}
audiotactile <- data[which(data$CONDITION == 1),]  
onlytactile <- data[which(data$CONDITION == 2),]

st.audiotactile <- shapiro.test(audiotactile$WPM)
st.onlytactile <- shapiro.test(onlytactile$WPM)

par(mfrow=c(1,2))
plotNormalHistogram(audiotactile$WPM, main="Audio Tactile")
plotNormalHistogram(onlytactile$WPM, main="Tactile")
```	

Upon first glance, it appears as though the data is normally distributed. However, p-values for shapiro.tests on the distributions are significant, indicating that the **data may not be normal**. The p-values for audio tactile and tactile are as follows: `r format(st.audiotactile$p.value,scientific=FALSE)` and `r format(st.onlytactile$p.value,scientific=FALSE)`.

By digging into the histograms further, we can see that the values are potentially **multimodal**. 

```{r}
par(mfrow=c(1,2))
hist(audiotactile$WPM, main="Audio Tactile", breaks=50)
hist(onlytactile$WPM, main="Tactile", breaks=50)
```	

This potential multimodal distribution can be likely attributed to the low sample size (n = `r length(unique(data$ID))`). With more samples, we could expect to see a more normal distribution. Nevertheless, we attempt to apply some transformations to see if we can normalize the data and apply parametric stats. 


We attempt the following transformations: **Square Root**, **Cube Root**, and **Log**. 
```{r}
# Square Root
st.at.sqrt <- shapiro.test(sqrt(audiotactile$WPM))
st.tt.sqrt <- shapiro.test(sqrt(onlytactile$WPM))

# Cube Root

st.at.cube <- shapiro.test(sign(audiotactile$WPM) * abs(audiotactile$WPM)^(1/3))
st.tt.cube <- shapiro.test(sign(onlytactile$WPM) * abs(onlytactile$WPM)^(1/3))

# Log
st.at.log <- shapiro.test(log(audiotactile$WPM))
st.tt.log <- shapiro.test(log(onlytactile$WPM))

```


